---
title: "Iterations and listcols"
output: github_document
date: "2025-10-28"
---

```{r}
library(tidyverse)
library(rvest)
```

note: `l[1]` means first item in your list
```{r}
l = 
  list(
    vec_numeric = 1:23,
    char_vec = c("chelsey"),
    mat = matrix(1:8, nrow = 2, ncol = 4),
    summary = summary(rnorm(1000, mean = 4))
  )

 l

l[1]

l[["vec_numeric"]]

```

## Make a different list
```{r}
list_normals = 
  list(
    a = rnorm(30, mean = 3, sd = 1),
    b = rnorm(30, mean = 30, sd = 1),
    c = rnorm(30, mean = 3, sd = 10),
    d = rnorm(30, mean = -3, sd = 4)
  )
```

(copy and paste the function from last time)
```{r}
mean_and_sd = function(x) {
  
 if(!is.numeric(x)) {
  stop("the input x shoud be numeric")
}
  
if(length(x) < 5) {
  stop("only compute mean and sd when the input has 5 or more numbers") 
}

mean_x = mean(x, na.rm = TRUE)
sd_x = sd(x, na.rm = TRUE)

tibble(
  mean = mean_x,
  sd = sd_x
)

}
```

```{r}
mean_and_sd(list_normals[[1]])
mean_and_sd(list_normals[[2]])
mean_and_sd(list_normals[[3]])
mean_and_sd(list_normals[[4]])
```

use a for loop to iterate!
```{r}
output = vector("list", length = 4)

for (i in 1:4) {
  
  output[[i]] = mean_and_sd(list_normals[[i]])
}

output
```

use `map` to do the same thing!
`list_normals` = this is the input we are iterating each time and we are using this = `mean_and_sd` as the function
```{r}
output = map(list_normals, mean_and_sd)

output
```

check out some `map` variants
note: `map_dfr` collapses output into dataframe
note: `.id` names all your output
note: `map_dbl(list_normals, median)` gives you the median of your outputs in one line
```{r}
map_dfr(list_normals, mean_and_sd, .id = "sample")

map_dbl(list_normals, median)
```

## List columns

try to put my list into a dataframe bc they have the same length
```{r}
listcol_df = 
  tibble(
    name = c("a", "b", "c", "d"),
    sample = list_normals
  )
```

did this really work?
```{r}
pull(listcol_df, sample)
pull(listcol_df, name)
```

can I apply `mean_and_sd`?
```{r}
mean_and_sd(pull(listcol_df, sample)[[1]])
mean_and_sd(pull(listcol_df, sample)[[2]])
mean_and_sd(pull(listcol_df, sample)[[3]])
mean_and_sd(pull(listcol_df, sample)[[4]])
```

iterate using `map`
```{r}
map(pull(listcol_df, sample), mean_and_sd)
```

adding a column...
```{r}
listcol_df = 
  listcol_df %>% 
  mutate(
    summary = map(sample, mean_and_sd)
  )

```

note: `unnest` means to expand table
```{r}
listcol_df %>% 
  select(-sample) %>% 
  unnest(summary)
```

## Revist NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

nsduh_import = function(html, table_num) {

data =
  html %>% 
  html_table() |> 
  nth(table_num) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))

data

}

nsduh_import(nsduh_html, table_num = 1)
nsduh_import(nsduh_html, table_num = 2)
nsduh_import(nsduh_html, table_num = 3)

```

try this with for loop
```{r}
output = vector("list", length = 3)

for (i in 1:3) {
  
  output[[1]] = nsduh_import(html = nsduh_html, i)
}
```

do this with map
```{r}
map(1:3, nsduh_import, html = nsduh_html)
```

do this all in a dataframe
```{r}
nsduh_df = 
  tibble(
    name = c("marj year", "marj month", "marj first"),
    number = 1:3
  ) %>% 
  mutate(
    table = map(number, nsduh_import, html = nsduh_html)
  ) %>% 
  unnest(table)
```

## look at weather data

```{r}
library(p8105.datasets)
data("weather_df")
```

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point() +
  geom_smooth(method = "lm")
```

let's do a regression
```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  lm(tmax ~ tmin, data = _)

weather_df %>% 
  filter(name == "Molokai_HI") %>% 
  lm(tmax ~ tmin, data = _)

weather_df %>% 
  filter(name == "Waterhole_WA") %>% 
  lm(tmax ~ tmin, data = _)
```

let's iterate differently
```{r}
weather_nest_df = weather_df %>% 
  nest(data = date:tmin)
```

```{r}
lm(tmax ~ tmin, data = pull(weather_nest_df, data)[[1]])
lm(tmax ~ tmin, data = pull(weather_nest_df, data)[[2]])
lm(tmax ~ tmin, data = pull(weather_nest_df, data)[[3]])
```

do this using map
```{r}
weather_lm = function(df) {
  
  lm(tmax ~ tmin, data = df)
}
```

```{r}
map(pull(weather_nest_df, data), weather_lm)
```


```{r}
weather_nest_df %>% 
  mutate(
    lm_fits = map(data, weather_lm)
  )
```

